---
title: "Statistical Inference part 1"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggpubr)
```

Final project for Statistical Inference by Jose Ramon Pineda

Packages used: tidyverse, ggpubr

#PART 1
```{r, part1}


lambda <-  0.2
n <- 40
sim <- 1000
set.seed(seed = 421)
mu <- 1/lambda
```

To begin, we expect a mean of 5 (1/lambda)

```{r, b}
test <- replicate(sim, rexp(n, lambda))
meantest <- data.frame(means = apply(test, 2, mean))

mean(meantest$means)

#Sample mean turns out to be close to the expected 5 at 5.02

var <- ((1/lambda)/sqrt(n))^2
var
testvar <- var(meantest$means)
testvar

#We see that the expected variance (6.25) is a bit higher than the variance we got in the test (5.99)

ggplot(data = meantest, aes(x = means)) + 
        geom_histogram(binwidth=0.1, aes(y=..density..), alpha=0.2) + 
        geom_vline(xintercept = mu, size=1, colour="red") + 
        geom_density(colour="blue", size=1) +
        scale_x_continuous(breaks=seq(mu-5,mu+5,1), limits=c(mu-5,mu+6))  +
        labs(title = "Is this a normal distribution?", subtitle = "At first glance, yes") +
        stat_function(fun = dnorm, args = list(mean = mu , sd = sqrt(var)), colour = "red", size=1)
        
```


Graphically this seems like a normal distribution. The blue line shows our test data and the red shows what a normal distribution would look like. Let's run a Shapiro test to confirm that it is actually normal.

```{r, c}
shapiro.test(meantest$means)
ggqqplot(meantest$means, title = "QQ plot")

```

The shapiro test shows an extremely small p-value, so we reject the Ho that this is normally distributed.

We also run a qqplot to visualize this some more. We see then that some of the points on the right tail fall outside the confidence interval testing normality, which likely explains the Shapiro test results.






